\chapter{Modal analysis}\label{chap:modalanalysis}
When external forces act on a multi degree-of-freedom system, the system
undergoes forced vibration.
For a system with three coordinates or degrees of freedom, the governing
equation of motion are a set of three coupled ordinary differential equations of
second second order.
The equations of motion of a multi degree-of-freedom system under external force
are given by:
\begin{equation}
  \label{eq:modal}
  [\mathbf{M}] \ddot{\vec{x}} + [\mathbf{K}] \vec{x} = \vec{\mathbf{F}}
\end{equation}
The equation \eqref{eq:modal} represents the system in the version without
damping, where \(\vec{\mathbf{F}}\) is the vector of the external force.
\section{Forced vibration of undamped system}\label{sec:eigen}
To solve the equation \eqref{eq:modal} by modal analysis, it is necessary first
to solve the eigenvalue problem and find the natural frequencies \(\omega_1,
\dots, \omega_n\) as reported below:
\(\omega_{1} = 8.27843\) [\si{\radian\per\second}],
\(\omega_{2} = 27.40935\) [\si{\radian\per\second}],
\(\omega_{3} = 41.87035\) [\si{\radian\per\second}];
and the corresponding normal modes:
\begin{equation}
	\label{eq:eiegnmodefree}
 	[\mathbf{U}] = \begin{bmatrix*}[r]
 		1.00000 & 1.00000 & 1.00000 \\
		0.86540 &-0.47555 &-2.44325 \\
		0.61945 &-1.28034 & 2.15533 \\
 	\end{bmatrix*}
\end{equation}
\section{Forced vibration of viscous damped system}\label{sec:viscoussystem}
Modal analysis, as presented in previous section \ref{sec:eigen}, applies only
to undamped systems.
In many cases, the influence of damping upon the response of a vibratory system
is minor and can be disregarded. However, it must be considered if the response
of the system is required for a relatively long period of time compared to the
natural periods of the system. Further, if the frequency of excitation is at or
near one of the natural frequencies of the system, damping is of primary
importance and must be taken into account. In general, since the effects are not
known in advance, damping must be considered in the vibration analysis of any
system, thus using the equation of motion:
\begin{equation}\label{eq:modaldamping}
  [\mathbf{M}] \ddot{\vec{x}} + [\mathbf{C}] \dot{\vec{x}} +[\mathbf{K}] \vec{x}
   = \vec{\mathbf{F}}
\end{equation}
%
Then we can consider for which the damping matrix can be expressed as a linear
combination of the mass and stiffness matrices:
\begin{equation}
\label{eq:moadlpropdamping}
	[C] = \alpha \cdot [\mathbf{M}] + \beta \cdot [\mathbf{K}]
\end{equation}
where \(\alpha\) and \(\beta\) are constant. By substituting in
\eqref{eq:modaldamping}, we obtain:
\begin{equation}\label{eq:modaldampingcomplete}
	[\mathbf{M}] \ddot{\vec{x}} +
	[\alpha [\mathbf{M}] + \beta[\mathbf{K}]] \dot{\vec{x}} +
	[\mathbf{K}] \vec{x}= \vec{\mathbf{F}}
\end{equation}
By expressing the solution vector \(\vec{x}\) as linear combination of the
natural mode of the undamped system. Substituting \(x(t) = [\mathbf{U}]q(t)\)
in the equation \eqref{eq:modaldampingcomplete} and pre-multiplying by 
\([\mathbf{U}]^{\top}\) we obtain:
\begin{equation}\label{eq:modaldampingexpansed}
	[\mathbf{U}]^{\top}[\mathbf{M}][\mathbf{U}] \ddot{\vec{q}} +
	[\alpha[\mathbf{U}]^{\top}[\mathbf{M}][\mathbf{U}] +
   \beta[\mathbf{U}]^{\top}[\mathbf{K}]][\mathbf{U}] \dot{\vec{q}} +
	[\mathbf{U}]^{\top}[\mathbf{K}][\mathbf{U}]^{\top} \vec{q} =
  [\mathbf{U}]^{\top}\vec{\mathbf{F}}
\end{equation}
%
If the eigenvectors \(\vec{\mathbf{U}}^{(j)}\) are normalized we can rewrite in 
decoupled form:
\begin{equation}\label{eq:decupled}
  \ddot{q}_{i}(t) + (\alpha + \omega_{i}^2\beta)\dot{q}_{i}(t) +
  \omega_{i}^{2}q(t) = Q_{i}(t)
\end{equation}
where: \(\omega_{1} = 8.27799\) [\si{\radian\per\second}],
\(\omega_{2} = 27.40863\) [\si{\radian\per\second}],
\(\omega_{3} = 41.91576\) [\si{\radian\per\second}] are
the natural frequencies of the undamped system.
The eigenvector are reported in \eqref{eq:eiegnmodeprop}.
\begin{equation}\label{eq:eiegnmodeprop}
	[\mathbf{U}] = \begin{bmatrix*}[r]
		1.00000 & 1.00000 & 1.00000 \\
		0.86499 &-0.48006 &-2.46147 \\
		0.61907 &-1.28528 & 2.16954 \\
 	\end{bmatrix*}
\end{equation}
In the equation \eqref{eq:decupled} substituting the relationship with 
\(\alpha + \omega^{2}_{i}\beta = 2\,\zeta_{i}\,\omega_{i}\), where
\(\zeta_{i}\) is called \emph{modal damping ratio} for the \(i_\text{th}\) normal
mode, equation \eqref{eq:decupled} we obtain:
\begin{equation}\label{eq:decupled2}
  \ddot{q}_{i}(t) + 2\zeta_{i}\omega_{i}\dot{q}_{i}(t) +
  \omega_{i}^{2}q(t) = Q_{i}(t)
\end{equation}
It can be seen that each of then equations represented by this expression is
uncoupled from all of the others.
Hence we can find the response of the \(i_\text{th}\) mode in the same manner
as that of a viscously damped single-degree-of-freedom system.
%
The identification of the sources and magnitude of damping is difficult in most
practical problems.
The condition given by equation \eqref{eq:moadlpropdamping} is sufficient but not
necessary for the existence of normal modes in damped systems.
The necessary condition is that the transformation that diagonalizes the damping
matrix also uncouples the coupled equations of motion.
This condition is less restrictive than equation \eqref{eq:moadlpropdamping}
and covers more possibilities.
In the general case of damping, the damping matrix cannot be diagonalized
simultaneously with the mass and stiffness matrices.
A common procedure for finding the solution of the eigenvalue problem of a
damped system involves the transformation of the coupled second-order
equations of motion into \(2n\) uncoupled first-order equation.
%
\input{rayleigh}
\input{matrixiterationmethod}
